---
title: "Environmental Enforcement as a Credit Allocation Mechanism"
subtitle: "A Staggered Difference-in-Differences Analysis"
author: "Josefina Silva Fuentealba"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    code-summary: "Show code"
    fig-width: 9
    fig-height: 6
    fig-dpi: 150
    embed-resources: true
    smooth-scroll: true
    highlight-style: github
execute:
  warning: false
  message: false
  echo: true
  cache: false
---

```{r setup}
#| label: setup
#| include: false

# ── Core packages ──────────────────────────────────────────────────────────────
library(haven)          # read .dta files
library(tidyverse)      # data wrangling + ggplot2
library(fixest)         # TWFE with high-dimensional FE (feols)
library(did)            # Callaway & Sant'Anna (2021)
library(bacondecomp)    # Goodman-Bacon decomposition
library(modelsummary)   # regression tables
library(kableExtra)     # table formatting
library(patchwork)      # combine ggplot panels
library(scales)         # axis formatting
library(gt)             # nice tables
library(DescTools)      # Winsorize
library(sessioninfo)    # session info

# ── Global ggplot theme ────────────────────────────────────────────────────────
theme_project <- function() {
  theme_minimal(base_size = 12) +
    theme(
      plot.title       = element_text(face = "bold", size = 13),
      plot.subtitle    = element_text(color = "grey40", size = 11),
      axis.title       = element_text(size = 11),
      panel.grid.minor = element_blank(),
      legend.position  = "bottom",
      plot.caption     = element_text(color = "grey50", size = 9)
    )
}
theme_set(theme_project())

# ── Project colour palette ─────────────────────────────────────────────────────
col_high  <- "#1B4F72"   # high ENCORE (navy)
col_low   <- "#E74C3C"   # low  ENCORE (red)
col_court <- "#27AE60"   # court introduction (green)

# ── Paths ──────────────────────────────────────────────────────────────────────
# Input data  (raw .dta files from CMF project)
path_data <- "C:/Users/wb631166/OneDrive - WBG/Desktop/CEMLA/CMF/Data"

# Output directory (figures, tables, processed datasets)
path_out  <- "C:/WBG/GitHub/CMF/Data/out"

# Create output directory if it does not exist
if (!dir.exists(path_out)) dir.create(path_out, recursive = TRUE)

# Helper: build full input path
din  <- function(...) file.path(path_data, ...)

# Helper: build full output path
dout <- function(...) file.path(path_out,  ...)
```

---

# Data Cleaning {#sec-cleaning}

## CMF Credit Registry

The CMF credit registry (`cmf_project_2011-2024.dta`) contains the universe of corporate loans issued by Chilean banks. We keep only firms matched with the SII tax registry (`mergeC11D03 == 3`), construct the **new-credit variable** as the positive first difference in outstanding loan balances at the bank–firm–month level, and collapse to annual frequency.

```{r load-cmf}
#| label: load-cmf

# ── 1. Load ───────────────────────────────────────────────────────────────────

cmf_raw <- read_dta(din("Base Final/cmf_project_2011-2024.dta"))

# ── 2. Keep matched firms only ────────────────────────────────────────────────
cmf <- cmf_raw |>
  filter(mergeC11D03 == 3) |>
  select(-mergeC11D03)

# ── 3. Parse date from periodo (YYYYMM integer) ───────────────────────────────
cmf <- cmf |>
  mutate(
    year  = as.integer(substr(as.character(periodo), 1, 4)),
    month = as.integer(substr(as.character(periodo), 5, 6))
  )

# ── 4. Identify new loans: positive Δ in outstanding balance ──────────────────
cmf <- cmf |>
  arrange(id, ins_cod, year, month) |>
  mutate(
    c11_cartera_normal = if_else(c11_cartera_normal == 0,
                                 NA_real_, c11_cartera_normal)
  ) |>
  group_by(id, ins_cod) |>
  mutate(
    delta_cred      = c11_cartera_normal - lag(c11_cartera_normal),
    newcred_cartera = if_else(delta_cred > 0, delta_cred, NA_real_)
  ) |>
  ungroup()

# ── 5. Collapse to bank–firm–year ─────────────────────────────────────────────
cmf_annual <- cmf |>
  filter(!is.na(newcred_cartera)) |>
  group_by(id, ins_cod, year, región, subrubroeconómico) |>
  summarise(
    newcred_cartera = sum(newcred_cartera, na.rm = TRUE),
    .groups = "drop"
  )

cat("Annual panel:", nrow(cmf_annual), "bank-firm-year obs\n")
cat("Unique firms:", n_distinct(cmf_annual$id), "\n")
cat("Years:       ", paste(range(cmf_annual$year), collapse = "-"), "\n")
```

## Regional IDs and Court Entry Dates

We assign numeric region IDs and code the staggered court entry year (`year_court`). Three environmental courts were established under Law No. 20,600:

| Court | Regions | Entry year |
|-------|---------|------------|
| Santiago | 5, 6, 7, 13, 16 | 2013 |
| Valdivia | 8, 9, 10, 11, 12, 14 | 2014 |
| Antofagasta | 1, 2, 3, 4, 15 | 2017 |

```{r region-ids}
#| label: region-ids

region_lookup <- tribble(
  ~región,                                                          ~regid,
  "I REGION DE TARAPACA",                                              1L,
  "II REGION DE ANTOFAGASTA",                                          2L,
  "III REGION DE ATACAMA",                                             3L,
  "IV REGION COQUIMBO",                                                4L,
  "V REGION VALPARAISO",                                               5L,
  "VI REGION DEL LIBERTADOR GENERAL BERNARDO O'HIGGINS",               6L,
  "VII REGION DEL MAULE",                                              7L,
  "VIII REGION DEL BIO BIO",                                           8L,
  "IX REGION DE LA ARAUCANIA",                                         9L,
  "X REGION LOS LAGOS",                                               10L,
  "XI REGION AYSEN DEL GENERAL CARLOS IBAÑEZ DEL CAMPO",             11L,
  "XII REGION DE MAGALLANES Y LA ANTARTICA CHILENA",                  12L,
  "XIII REGION METROPOLITANA",                                        13L,
  "XIV REGION DE LOS RIOS",                                           14L,
  "XV REGION ARICA Y PARINACOTA",                                     15L,
  "XVI REGION DE ÑUBLE",                                              16L
)

court_lookup <- tribble(
  ~regid, ~year_court,
  # Santiago court — January 2013
   5L, 2013L,  6L, 2013L,  7L, 2013L, 13L, 2013L, 16L, 2013L,
  # Valdivia court — December 2013 (first full year = 2014)
   8L, 2014L,  9L, 2014L, 10L, 2014L, 11L, 2014L, 12L, 2014L, 14L, 2014L,
  # Antofagasta court — July 2017
   1L, 2017L,  2L, 2017L,  3L, 2017L,  4L, 2017L, 15L, 2017L
)

cmf_annual <- cmf_annual |>
  left_join(region_lookup, by = "región") |>
  left_join(court_lookup,  by = "regid") |>
  mutate(
    # never-treated = 0 (required by the did package)
    year_court = replace_na(year_court, 0L),
    court      = as.integer(year_court > 0 & year >= year_court)
  )
```

## SMA Enforcement Cases

```{r sma-cases}
#| label: sma-cases

sma_raw <- read_dta(din("cmf_project_super_intendencia.dta")) |>
  mutate(year = as.integer(year))  # viene como string, hay que convertir

# Construir regid desde region (string) usando el mismo lookup de arriba
region_lookup_sma <- tribble(
  ~region,                                                              ~regid,
  "Región de Tarapacá",                                                     1L,
  "Región de Antofagasta",                                                   2L,
  "Región de Atacama",                                                       3L,
  "Región de Coquimbo",                                                      4L,
  "Región de Valparaíso",                                                    5L,
  "Región del Libertador General Bernardo O'Higgins",                        6L,
  "Región del Maule",                                                        7L,
  "Región del Biobío",                                                       8L,
  "Región de la Araucanía",                                                  9L,
  "Región de los Lagos",                                                    10L,
  "Región de Aysén del General Carlos Ibáñez del Campo",                   11L,
  "Región de Magallanes y la Antártica Chilena",                           12L,
  "Región Metropolitana",                                                   13L,
  "Región de Los Ríos",                                                     14L,
  "Región de Arica y Parinacota",                                           15L,
  "Región de Ñuble",                                                        16L
)

# Contar casos por region-year y calcular lrcases (lag log)
reg_cases <- sma_raw |>
  left_join(region_lookup_sma, by = "region") |>
  filter(!is.na(regid)) |>
  group_by(regid, year) |>
  summarise(cases = n(), .groups = "drop") |>
  arrange(regid, year) |>
  group_by(regid) |>
  mutate(lrcases = lag(log1p(cases))) |>   # one-year lag, as in paper
  ungroup()

# Merge into annual panel
cmf_annual <- cmf_annual |>
  left_join(reg_cases |> select(regid, year, lrcases),
            by = c("regid", "year")) |>
  mutate(lrcases = replace_na(lrcases, 0))

cat("SMA cases merged. Regions covered:", n_distinct(reg_cases$regid), "\n")
cat("Years covered:", paste(range(reg_cases$year, na.rm = TRUE), collapse = "-"), "\n")
```


## ENCORE Environmental Impact Scores

```{r encore}
#| label: encore

# super_IDs.dta is already built by your Stata pipeline (CR_WDD_mob.do)
encore <- read_dta(din("super_IDs.dta")) |>
  select(subrubroeconómico, score_share, d50_score, d75_score) |>
  mutate(subrubroeconómico = str_trim(subrubroeconómico))

cmf_annual <- cmf_annual |>
  mutate(subrubroeconómico = str_trim(subrubroeconómico)) |>
  left_join(encore, by = "subrubroeconómico")
```

## Final Working Sample

```{r working-sample}
#| label: working-sample

df <- cmf_annual |>
  # Drop Santiago (as in paper) and rows without region / ENCORE info
  filter(regid != 13, !is.na(regid), !is.na(score_share)) |>
  mutate(
    lnewcred   = log(newcred_cartera),
    # Winsorize at 1% — DescTools syntax uses quantile, not probs
    lnewcred_w = Winsorize(lnewcred,
                           quantile(lnewcred, probs = c(0.01, 0.99),
                                    na.rm = TRUE)),
    # Integer IDs for FE
    bankfirm_id = as.integer(interaction(ins_cod, id, drop = TRUE)),
    IDB         = as.integer(factor(ins_cod)),
    IDR         = as.integer(factor(regid)),
    IDS         = as.integer(factor(subrubroeconómico))
  )

cat("Final sample:\n")
cat("  Obs:             ", nrow(df), "\n")
cat("  Bank-firm pairs: ", n_distinct(df$bankfirm_id), "\n")
cat("  Unique firms:    ", n_distinct(df$id), "\n")
cat("  Years:           ", paste(range(df$year), collapse = "-"), "\n")
cat("  Regions:         ", n_distinct(df$regid), "\n")
cat("\nCohort distribution (bank-firm level):\n")
df |> distinct(bankfirm_id, year_court) |> count(year_court) |> print()

# ── Save cleaned working sample ────────────────────────────────────────────────
write_rds(df, dout("df_working_sample.rds"))
write_csv(df, dout("df_working_sample.csv"))
cat("\nWorking sample saved to:", path_out, "\n")
```


```{r never-treated}
#| label: never-treated


# ── Define never-treated following the paper's approach ───────────────────────
# Firms in sectors with score_share < 0.2 are assigned year_court = 0
# (never-treated) because environmental courts are unlikely to affect
# sectors with negligible environmental impact.
df <- df |>
  mutate(
    year_court = if_else(score_share < 0.2, 0L, year_court),
    court      = if_else(score_share < 0.2, 0L, court)
  )

# ── Fix: year_court must be time-invariant per bankfirm_id ────────────────────
# Some pairs change region or score_share across years (~2% of sample).
# We assign the modal (most frequent) year_court value per pair.
year_court_fixed <- df |>
  group_by(bankfirm_id) |>
  summarise(
    year_court = as.integer(names(which.max(table(year_court)))),
    .groups = "drop"
  )

df <- df |>
  select(-year_court) |>
  left_join(year_court_fixed, by = "bankfirm_id")

# ── Verify cohort distribution ─────────────────────────────────────────────────
cat("Cohort distribution after never-treated assignment:\n")
df |> distinct(bankfirm_id, year_court) |> count(year_court) |> print()

cat("\nNot-yet-treated + never-treated by year:\n")
df |>
  mutate(nyt = as.integer(year < year_court | year_court == 0)) |>
  group_by(year) |>
  summarise(
    treated     = sum(court == 1),
    control     = sum(nyt == 1),
    pct_control = round(100 * mean(nyt == 1), 1),
    .groups     = "drop"
  ) |>
  print(n = 14)

# ── Build df_cs ready for Callaway-Sant'Anna ───────────────────────────────────
df_cs <- df |>
  filter(!is.na(lnewcred_w), !is.na(lrcases), !is.na(score_share)) |>
  mutate(year_court = as.integer(year_court))

# Final check: no pair should have more than one year_court value
inconsistent <- df_cs |>
  group_by(bankfirm_id) |>
  summarise(n_values = n_distinct(year_court), .groups = "drop") |>
  filter(n_values > 1) |>
  nrow()

cat("\nPares con year_court inconsistente después del fix:", inconsistent, "\n")
cat("df_cs listo para CS —", nrow(df_cs), "obs |",
    n_distinct(df_cs$bankfirm_id), "bank-firm pairs\n")
```



---

# Descriptive Statistics {#sec-descriptives}

## Summary Table

```{r summary-table}
#| label: summary-table

vars_summary <- df |>
  select(
    `Log new credit (winsorised)` = lnewcred_w,
    `Log SMA cases (lag)`         = lrcases,
    `ENCORE score`                = score_share,
    `Court (binary)`              = court
  )

datasummary_skim(
  vars_summary,
  title  = "Table 1: Summary Statistics",
  notes  = "Sample: bank-firm-year observations 2011-2024. Santiago excluded.",
  output = "kableExtra"
) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## New Credit Over Time by ENCORE Group

```{r credit-time}
#| label: credit-time
#| fig-cap: "Aggregate new credit by ENCORE exposure group (2011-2024). Dashed lines mark court introduction years."

credit_ts <- df |>
  filter(!is.na(d50_score)) |>
  group_by(year, d50_score) |>
  summarise(total = sum(exp(lnewcred_w), na.rm = TRUE) / 1e9, .groups = "drop") |>
  mutate(group = if_else(d50_score == 1, "High ENCORE", "Low ENCORE"))

ymax <- max(credit_ts$total, na.rm = TRUE)

ggplot(credit_ts, aes(year, total, colour = group)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2) +
  geom_vline(xintercept = c(2013, 2014, 2017),
             linetype = "dashed", alpha = 0.5, colour = col_court) +
  annotate("text", x = c(2013.15, 2014.15, 2017.15),
           y = ymax * c(0.95, 0.85, 0.75),
           label = c("Santiago\ncourt", "Valdivia\ncourt", "Antofagasta\ncourt"),
           size = 3, hjust = 0, colour = col_court) +
  scale_colour_manual(values = c(col_high, col_low)) +
  scale_y_continuous(labels = label_number(suffix = "B CLP")) +
  scale_x_continuous(breaks = 2011:2024) +
  labs(
    title    = "New Credit by ENCORE Exposure Group",
    subtitle = "High ENCORE = sectors above median ENCORE score",
    x = NULL, y = "Aggregate new credit (billions CLP)",
    colour = NULL
  )
ggsave(dout("fig1_credit_by_encore.png"), width = 9, height = 6, dpi = 150)
```

## SMA Enforcement Activity

```{r sma-plots}
#| label: sma-plots
#| fig-cap: "Left: total SMA cases per year. Right: cumulative cases by region."

p1 <- sma_raw |>
  count(year) |>
  filter(year >= 2011) |>
  ggplot(aes(year, n)) +
  geom_col(fill = col_high, alpha = 0.85) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2011, 2024, 2)) +
  labs(title = "SMA Cases per Year", x = NULL, y = "Cases")

p2 <- sma_raw |>
  count(region, name = "cases") |>
  slice_max(cases, n = 12) |>
  mutate(region = str_wrap(region, 25),
         region = fct_reorder(region, cases)) |>
  ggplot(aes(cases, region)) +
  geom_col(fill = col_high, alpha = 0.85) +
  scale_x_continuous(labels = comma) +
  labs(title = "Total Cases by Region (Top 12)", x = "Cases", y = NULL)

p_sma <- p1 + p2
p_sma
ggsave(dout("fig2_sma_activity.png"), plot = p_sma, width = 9, height = 5, dpi = 150)
```

## ENCORE Score Distribution

```{r encore-dist}
#| label: encore-dist
#| fig-cap: "Distribution of ENCORE scores across sectors. Dashed line = median."

p_enc <- df |>
  distinct(subrubroeconómico, score_share) |>
  ggplot(aes(score_share)) +
  geom_histogram(bins = 20, fill = col_high, alpha = 0.85, colour = "white") +
  geom_vline(xintercept = median(df$score_share, na.rm = TRUE),
             linetype = "dashed", colour = col_low, linewidth = 1) +
  annotate("text",
           x = median(df$score_share, na.rm = TRUE) + 0.01, y = 3.8,
           label = "Median", colour = col_low, hjust = 0, size = 3.5) +
  labs(title = "ENCORE Score by Sector",
       x = "ENCORE score (normalised 0-1)", y = "Count of sectors")

p_enc
ggsave(dout("fig3_encore_dist.png"), plot = p_enc, width = 7, height = 5, dpi = 150)
```

---

# Baseline TWFE Replication {#sec-twfe}

We replicate the paper's two main specifications using `fixest::feols()`.

## Equation 1 — SMA Cases × ENCORE Score

$$\text{Loan}_{i,j,r,s,t} = \alpha_0 + \beta_1 \text{Cases}_{r,t-1} + \beta_2 \text{Score}_s + \beta_3(\text{Cases}_{r,t-1} \times \text{Score}_s) + \mu_{ij} + \delta_{st} + \omega_{jt} + u$$

```{r twfe-eq1}
#| label: twfe-eq1

df_high <- df |>
  filter(score_share >= quantile(score_share, 0.25, na.rm = TRUE))

# Table 2 replication: cases only, high-ENCORE sample
m2_1 <- feols(lnewcred_w ~ log1p(lrcases) | bankfirm_id,
              data = df_high, cluster = ~IDR + IDS)
m2_2 <- feols(lnewcred_w ~ log1p(lrcases) | bankfirm_id + IDS^year,
              data = df_high, cluster = ~IDR + IDS)
m2_3 <- feols(lnewcred_w ~ log1p(lrcases) | bankfirm_id + IDS^year + IDB^year,
              data = df_high, cluster = ~IDR + IDS)

# Table 3 replication: interaction with continuous ENCORE score
m3_1 <- feols(lnewcred_w ~ log1p(lrcases) * score_share | bankfirm_id,
              data = df, cluster = ~IDR + IDS)
m3_2 <- feols(lnewcred_w ~ log1p(lrcases) * score_share | bankfirm_id + IDS^year,
              data = df, cluster = ~IDR + IDS)
m3_3 <- feols(lnewcred_w ~ log1p(lrcases) * score_share |
                bankfirm_id + IDS^year + IDB^year,
              data = df, cluster = ~IDR + IDS)

modelsummary(
  list("(1)" = m2_1, "(2)" = m2_2, "(3)" = m2_3,
       "(4)" = m3_1, "(5)" = m3_2, "(6)" = m3_3),
  title    = "Tables 2-3 Replication: Effect of SMA Cases on New Credit",
  stars    = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = c(
    "log1p(lrcases)"             = "Log Cases (lag)",
    "score_share"                = "ENCORE score",
    "log1p(lrcases):score_share" = "Log Cases × ENCORE"
  ),
  gof_map  = c("nobs", "r.squared"),
  notes    = "Clustered SE by region × sector. Cols 1-3: above-p25 ENCORE sample.",
  output   = "kableExtra"
) |>
  add_header_above(c(" " = 1,
                     "Table 2: Cases only (above-p25 ENCORE)" = 3,
                     "Table 3: Cases × ENCORE (full sample)" = 3)) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

## Equation 2 — Court Introduction × SMA Cases

$$\text{Loan}_{i,j,r,s,t} = \alpha_0 + \beta_1 \text{Cases}_{r,t-1} + \beta_2 \text{Court}_{r,t} + \beta_3(\text{Cases}_{r,t-1} \times \text{Court}_{r,t}) + \mu_{ij} + \delta_{st} + \omega_{jt} + u$$

The key coefficient is $\beta_3$: it captures whether judicial enforceability amplifies the credit response to regulatory pressure.

```{r twfe-eq2}
#| label: twfe-eq2

m4_1 <- feols(lnewcred_w ~ log1p(lrcases) * court | bankfirm_id + IDS^year,
              data = df, cluster = ~IDR + IDS)
m4_2 <- feols(lnewcred_w ~ log1p(lrcases) * court |
                bankfirm_id + IDS^year + IDB^year,
              data = df, cluster = ~IDR + IDS)
m4_3 <- feols(lnewcred_w ~ log1p(lrcases) * court |
                bankfirm_id + IDS^year + IDB^year,
              data = df |> filter(d50_score == 1), cluster = ~IDR + IDS)
m4_4 <- feols(lnewcred_w ~ log1p(lrcases) * court |
                bankfirm_id + IDS^year + IDB^year,
              data = df |> filter(d50_score == 0), cluster = ~IDR + IDS)

modelsummary(
  list("Full (1)" = m4_1, "Full (2)" = m4_2,
       "High ENCORE" = m4_3, "Low ENCORE" = m4_4),
  title    = "Table 4 Replication: Court Introduction × SMA Cases",
  stars    = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = c(
    "log1p(lrcases)"       = "Log Cases (lag)",
    "court"                = "Court",
    "log1p(lrcases):court" = "Log Cases × Court"
  ),
  gof_map  = c("nobs", "r.squared"),
  notes    = "Clustered SE by region × sector. Santiago excluded.",
  output   = "kableExtra"
) |>
  add_header_above(c(" " = 1, "Full sample" = 2, "By ENCORE exposure" = 2)) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

---

# Callaway–Sant'Anna Staggered DiD {#sec-cs}

Standard TWFE can be biased when treatment effects are heterogeneous across cohorts, because it implicitly uses already-treated units as controls for later-treated units. The Callaway & Sant'Anna (2021) estimator avoids this by constructing clean 2×2 ATT(g,t) comparisons for each cohort $g$ and period $t$.

**Setup:**

- **Treatment:** `year_court` — year the court became operational in the firm's region (0 = never-treated)
- **Unit:** `bankfirm_id` (bank–firm pair)
- **Controls:** `lrcases` + `score_share` (to improve parallel trends plausibility)
- **Comparison group:** not-yet-treated units (`control_group = "notyettreated"`)
- **Method:** doubly-robust (`est_method = "dr"`)

```{r cs-setup}
#| label: cs-setup

df_cs <- df |>
  filter(!is.na(lnewcred_w), !is.na(lrcases), !is.na(score_share)) |>
  mutate(year_court = as.integer(year_court))

cat("CS sample:", nrow(df_cs), "obs |",
    n_distinct(df_cs$bankfirm_id), "bank-firm pairs\n\n")
cat("Cohort distribution (bank-firm pairs):\n")
df_cs |>
  distinct(bankfirm_id, year_court) |>
  count(year_court, name = "pairs") |>
  mutate(label = case_when(
    year_court == 0    ~ "Never treated",
    year_court == 2013 ~ "Santiago (g=2013)",
    year_court == 2014 ~ "Valdivia (g=2014)",
    year_court == 2017 ~ "Antofagasta (g=2017)"
  )) |>
  select(label, pairs) |>
  print()
```

## Main CS Estimates — Full Sample

```{r cs-main}
#| label: cs-main
#| cache: true

# ── Correr CS sin covariables ──────────────────────────────────────────────────
# lrcases es constante dentro de cada región-año → causa matriz singular.
# Corremos sin xformla primero para verificar identificación base.
cs_main <- att_gt(
  yname         = "lnewcred_w",
  tname         = "year",
  idname        = "rs_id",
  gname         = "year_court",
  xformla       = ~ 1,              # sin covariables
  data          = df_rs_bal,
  control_group = "notyettreated",
  est_method    = "reg",
  clustervars   = "regid_fixed",
  print_details = FALSE
)

summary(cs_main)
```

## Aggregated ATT

```{r cs-agg}
#| label: cs-agg

cs_simple   <- aggte(cs_main, type = "simple")
cs_group    <- aggte(cs_main, type = "group")
cs_dynamic  <- aggte(cs_main, type = "dynamic", min_e = -3, max_e = 8)
cs_calendar <- aggte(cs_main, type = "calendar")

cat("─── Simple ATT ────────────────────────────────────────────────\n")
cat(sprintf("  ATT = %.4f  |  SE = %.4f  |  95%% CI: [%.4f, %.4f]\n",
    cs_simple$overall.att,
    cs_simple$overall.se,
    cs_simple$overall.att - 1.96 * cs_simple$overall.se,
    cs_simple$overall.att + 1.96 * cs_simple$overall.se))

cat("\n─── ATT by cohort ─────────────────────────────────────────────\n")
tibble(
  cohort  = cs_group$egt,
  ATT     = round(cs_group$att.egt, 4),
  SE      = round(cs_group$se.egt,  4)
) |>
  mutate(`95% CI` = sprintf("[%.4f, %.4f]",
                             ATT - 1.96 * SE, ATT + 1.96 * SE)) |>
  print()
```

## CS by ENCORE Subgroup

```{r cs-subgroups}
#| label: cs-subgroups
#| cache: true

# ── High ENCORE ────────────────────────────────────────────────────────────────
cs_high <- att_gt(
  yname         = "lnewcred_w",
  tname         = "year",
  idname        = "rs_id",
  gname         = "year_court",
  xformla       = ~ 1,
  data          = df_rs_bal |> filter(score_share >= median(score_share, na.rm = TRUE)),
  control_group = "notyettreated",
  est_method    = "reg",
  clustervars   = "regid_fixed",
  print_details = FALSE
)

# ── Low ENCORE ─────────────────────────────────────────────────────────────────
cs_low <- att_gt(
  yname         = "lnewcred_w",
  tname         = "year",
  idname        = "rs_id",
  gname         = "year_court",
  xformla       = ~ 1,
  data          = df_rs_bal |> filter(score_share < median(score_share, na.rm = TRUE)),
  control_group = "notyettreated",
  est_method    = "reg",
  clustervars   = "regid_fixed",
  print_details = FALSE
)

# ── Aggregate ──────────────────────────────────────────────────────────────────
cs_high_dyn <- aggte(cs_high, type = "dynamic", min_e = -3, max_e = 6)
cs_low_dyn  <- aggte(cs_low,  type = "dynamic", min_e = -3, max_e = 6)

cat("High ENCORE — simple ATT:", 
    round(aggte(cs_high, "simple")$overall.att, 4), "\n")
cat("Low  ENCORE — simple ATT:", 
    round(aggte(cs_low,  "simple")$overall.att, 4), "\n")
```

```{r cs-enforcement}
#| label: cs-enforcement
#| cache: true

# ── Clasificar regiones por intensidad de enforcement pre-tratamiento ──────────
# Usamos el promedio de lrcases ANTES de la entrada del tribunal
# para evitar endogeneidad — la intensidad es pre-determinada
enforcement_intensity <- df_rs_bal |>
  filter(year < year_court | year_court == 0) |>
  group_by(regid_fixed) |>
  summarise(mean_cases = mean(lrcases, na.rm = TRUE), .groups = "drop") |>
  mutate(
    high_enforce = as.integer(mean_cases >= median(mean_cases, na.rm = TRUE))
  )

cat("Regiones por intensidad de enforcement:\n")
enforcement_intensity |>
  mutate(label = if_else(high_enforce == 1, "High enforcement", "Low enforcement")) |>
  count(label) |> print()

# Merge back
df_rs_bal <- df_rs_bal |>
  left_join(enforcement_intensity |> select(regid_fixed, mean_cases, high_enforce),
            by = "regid_fixed")

# ── CS: regiones con enforcement ALTO ─────────────────────────────────────────
cs_high_enforce <- att_gt(
  yname         = "lnewcred_w",
  tname         = "year",
  idname        = "rs_id",
  gname         = "year_court",
  xformla       = ~ 1,
  data          = df_rs_bal |> filter(high_enforce == 1),
  control_group = "notyettreated",
  est_method    = "reg",
  clustervars   = "regid_fixed",
  print_details = FALSE
)

# ── CS: regiones con enforcement BAJO ─────────────────────────────────────────
cs_low_enforce <- att_gt(
  yname         = "lnewcred_w",
  tname         = "year",
  idname        = "rs_id",
  gname         = "year_court",
  xformla       = ~ 1,
  data          = df_rs_bal |> filter(high_enforce == 0),
  control_group = "notyettreated",
  est_method    = "reg",
  clustervars   = "regid_fixed",
  print_details = FALSE
)

# ── Aggregate ──────────────────────────────────────────────────────────────────
cs_he_simple <- aggte(cs_high_enforce, type = "simple")
cs_le_simple <- aggte(cs_low_enforce,  type = "simple")

cs_he_dyn <- aggte(cs_high_enforce, type = "dynamic", min_e = -2, max_e = 2)
cs_le_dyn <- aggte(cs_low_enforce,  type = "dynamic", min_e = -2, max_e = 2)

cat("\nSimple ATT by enforcement intensity:\n")
tibble(
  Group = c("High enforcement", "Low enforcement"),
  ATT   = round(c(cs_he_simple$overall.att, cs_le_simple$overall.att), 4),
  SE    = round(c(cs_he_simple$overall.se,  cs_le_simple$overall.se),  4)
) |>
  mutate(`95% CI` = sprintf("[%.3f, %.3f]",
                             ATT - 1.96 * SE,
                             ATT + 1.96 * SE)) |>
  print()

# ── Event study plot ───────────────────────────────────────────────────────────
tidy_cs <- function(cs_dyn, label) {
  tibble(
    event_time = cs_dyn$egt,
    att        = cs_dyn$att.egt,
    se         = cs_dyn$se.egt,
    group      = label
  ) |>
    mutate(ci_lo = att - 1.96 * se,
           ci_hi = att + 1.96 * se)
}

es_enforce <- bind_rows(
  tidy_cs(cs_he_dyn, "High enforcement"),
  tidy_cs(cs_le_dyn, "Low enforcement")
)

ggplot(es_enforce, aes(event_time, att, colour = group, fill = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = -0.5, linetype = "dotted", colour = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_colour_manual(values = c(col_high, col_low)) +
  scale_fill_manual(values   = c(col_high, col_low)) +
  scale_x_continuous(breaks  = -2:2) +
  facet_wrap(~group, ncol = 2) +
  labs(
    title    = "CS Event Study by Pre-Treatment Enforcement Intensity",
    subtitle = "Replicates Court × Cases logic using region-level enforcement stratification",
    x        = "Years relative to court introduction",
    y        = "ATT (log new credit)",
    colour   = NULL, fill = NULL
  ) +
  theme(legend.position = "none")

ggsave(dout("fig_cs_enforcement.png"), width = 10, height = 5, dpi = 150)
```


---

# Wooldridge ETWFE — Robustness {#sec-wooldridge}

As an additional robustness check we implement the Wooldridge (2021) **Extended TWFE (ETWFE)** estimator. This saturates interactions between cohort dummies, period dummies, and the continuous enforcement variable (`lrcases`), avoiding negative weighting while remaining in a familiar regression framework.

```{r etwfe}
#| label: etwfe


df_w <- df |>
  mutate(
    d2013 = as.integer(year_court == 2013),
    d2014 = as.integer(year_court == 2014),
    d2017 = as.integer(year_court == 2017),
    dinf  = as.integer(year_court == 0),
    f2013 = as.integer(year == 2013),
    f2014 = as.integer(year == 2014),
    f2015 = as.integer(year == 2015),
    f2016 = as.integer(year == 2016),
    f2017 = as.integer(year == 2017),
    f2018 = as.integer(year == 2018),
    f2019 = as.integer(year == 2019),
    f2020 = as.integer(year == 2020),
    # Time-invariant region ID for clustering
    IDR_fixed = as.integer(factor(regid))
  )

# ── Mean-centre lrcases within each cohort (correct grouping) ─────────────────
mean_rc <- df_w |>
  group_by(year_court) |>
  summarise(mean_lrcases = mean(lrcases, na.rm = TRUE), .groups = "drop")

df_w <- df_w |>
  left_join(mean_rc, by = "year_court") |>
  mutate(
    rc_dm13 = if_else(d2013 == 1, lrcases - mean_lrcases, 0),
    rc_dm14 = if_else(d2014 == 1, lrcases - mean_lrcases, 0),
    rc_dm17 = if_else(d2017 == 1, lrcases - mean_lrcases, 0)
  ) |>
  select(-mean_lrcases)

# ── ETWFE: saturated cohort × period × cases interactions ─────────────────────
etwfe <- feols(
  lnewcred_w ~
    # Cohort 2013: post-treatment periods
    I(d2013 * rc_dm13 * f2013) + I(d2013 * rc_dm13 * f2014) +
    I(d2013 * rc_dm13 * f2015) + I(d2013 * rc_dm13 * f2016) +
    I(d2013 * rc_dm13 * f2017) + I(d2013 * rc_dm13 * f2018) +
    I(d2013 * rc_dm13 * f2019) + I(d2013 * rc_dm13 * f2020) +
    # Cohort 2014: post-treatment periods
    I(d2014 * rc_dm14 * f2014) + I(d2014 * rc_dm14 * f2015) +
    I(d2014 * rc_dm14 * f2016) + I(d2014 * rc_dm14 * f2017) +
    I(d2014 * rc_dm14 * f2018) + I(d2014 * rc_dm14 * f2019) +
    I(d2014 * rc_dm14 * f2020) +
    # Cohort 2017: post-treatment periods
    I(d2017 * rc_dm17 * f2017) + I(d2017 * rc_dm17 * f2018) +
    I(d2017 * rc_dm17 * f2019) + I(d2017 * rc_dm17 * f2020) |
    bankfirm_id + IDS^year + IDB^year,
  data    = df_w,
  cluster = ~IDR_fixed + IDS
)

# ── Compute weighted average ATT across cohort-period cells ───────────────────
etwfe_coefs <- coef(etwfe)
etwfe_coefs <- etwfe_coefs[!is.na(etwfe_coefs)]
cat("ETWFE — average marginal effect (unweighted mean of cohort-period coefs):\n")
cat("  Mean =", round(mean(etwfe_coefs), 4), "\n")
cat("  Range: [", round(min(etwfe_coefs), 4), ",",
                  round(max(etwfe_coefs), 4), "]\n\n")

# ── Table ──────────────────────────────────────────────────────────────────────
modelsummary(
  list("ETWFE (Wooldridge 2021)" = etwfe),
  title   = "Robustness: Wooldridge ETWFE with Continuous Treatment",
  stars   = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  gof_map = c("nobs", "r.squared"),
  notes   = paste(
    "Each coefficient is a cohort × period × SMA cases interaction.",
    "lrcases is mean-centred within each treatment cohort.",
    "Clustered SE by region × sector."
  ),
  output  = "kableExtra"
) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

---

# Event Study Plots {#sec-event-study}



## Callaway–Sant'Anna Event Study

```{r event-cs}
#| label: event-cs
#| fig-cap: "CS event study by ENCORE subgroup. Shaded area = 95% CI. Solid points = statistically significant at 5%."

tidy_cs <- function(cs_dyn, label) {
  tibble(
    event_time = cs_dyn$egt,
    att        = cs_dyn$att.egt,
    se         = cs_dyn$se.egt,
    group      = label
  ) |>
    mutate(
      ci_lo = att - 1.96 * se,
      ci_hi = att + 1.96 * se,
      sig   = factor(if_else(ci_lo > 0 | ci_hi < 0, "p<0.05", "n.s."),
                     levels = c("p<0.05", "n.s."))
    )
}

es_data <- bind_rows(
  tidy_cs(cs_high_dyn, "High ENCORE"),
  tidy_cs(cs_low_dyn,  "Low ENCORE")
)

# Breaks dinámicos basados en los event times disponibles
et_range <- range(es_data$event_time, na.rm = TRUE)

ggplot(es_data, aes(event_time, att, colour = group, fill = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = -0.5, linetype = "dotted", colour = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(aes(shape = sig), size = 2.8) +
  scale_colour_manual(values = c(col_high, col_low)) +
  scale_fill_manual(values   = c(col_high, col_low)) +
  scale_shape_manual(values  = c("p<0.05" = 19, "n.s." = 1)) +
  scale_x_continuous(breaks  = seq(et_range[1], et_range[2], by = 1)) +
  facet_wrap(~group, ncol = 2) +
  labs(
    title    = "Callaway-Sant'Anna Event Study by ENCORE Subgroup",
    subtitle = "Panel A: High ENCORE  |  Panel B: Low ENCORE",
    x        = "Years relative to court introduction",
    y        = "ATT (log new credit)",
    colour   = NULL, fill = NULL, shape = "Significance"
  ) +
  theme(legend.position = "bottom")

ggsave(dout("fig4_cs_event_study_encore.png"), width = 10, height = 6, dpi = 150)
```

## ATT(g, t) by Treatment Cohort

```{r event-cohort}
#| label: event-cohort
#| fig-cap: "ATT(g,t) estimates for each cohort across calendar years. Dotted lines mark the year of court introduction."

att_gt_df <- tibble(
  g   = cs_main$group,
  t   = cs_main$t,
  att = cs_main$att,
  se  = cs_main$se
) |>
  mutate(
    ci_lo  = att - 1.96 * se,
    ci_hi  = att + 1.96 * se,
    sig    = factor(if_else(ci_lo > 0 | ci_hi < 0, "p<0.05", "n.s."),
                    levels = c("p<0.05", "n.s.")),
    # Labels dinámicos — solo incluye cohortes que realmente existen
    cohort = factor(g,
      levels = sort(unique(g)),
      labels = case_when(
        sort(unique(g)) == 2013 ~ "g=2013  Santiago",
        sort(unique(g)) == 2014 ~ "g=2014  Valdivia",
        sort(unique(g)) == 2017 ~ "g=2017  Antofagasta",
        TRUE                    ~ paste0("g=", sort(unique(g)))
      ))
  )

cat("Cohorts in cs_main:", paste(sort(unique(att_gt_df$g)), collapse = ", "), "\n")
cat("Calendar years covered:", paste(range(att_gt_df$t), collapse = "-"), "\n")

ggplot(att_gt_df, aes(t, att)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_vline(aes(xintercept = g - 0.5), linetype = "dotted",
             colour = col_court, alpha = 0.8) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi),
              fill = col_high, alpha = 0.15) +
  geom_line(colour = col_high, linewidth = 0.9) +
  geom_point(aes(shape = sig), colour = col_high, size = 2.5) +
  scale_shape_manual(values = c("p<0.05" = 19, "n.s." = 1)) +
  facet_wrap(~cohort, ncol = 2, scales = "free_x") +
  scale_x_continuous(breaks = seq(2011, 2024, 2)) +
  labs(
    title    = "ATT(g, t) by Treatment Cohort",
    subtitle = "Green dotted line = year of court introduction | solid points = p<0.05",
    x        = "Calendar year",
    y        = "ATT(g, t)",
    shape    = NULL
  ) +
  theme(legend.position = "bottom")

ggsave(dout("fig5_att_gt_cohorts.png"), width = 10, height = 5, dpi = 150)
```

---

# Heterogeneity Analysis {#sec-heterogeneity}

## By ENCORE Exposure and Cohort

```{r het-cohort}
#| label: het-cohort
#| fig-cap: "ATT by cohort for High vs Low ENCORE firms. Error bars = 95% CI."

tidy_group <- function(cs_obj, label) {
  g_agg <- aggte(cs_obj, type = "group")
  tibble(
    g      = g_agg$egt,
    att    = g_agg$att.egt,
    se     = g_agg$se.egt,
    group  = label
  ) |>
    mutate(
      ci_lo  = att - 1.96 * se,
      ci_hi  = att + 1.96 * se,
      sig    = factor(if_else(ci_lo > 0 | ci_hi < 0, "p<0.05", "n.s."),
                      levels = c("p<0.05", "n.s.")),
      # Labels dinámicos
      cohort = factor(g,
        levels = sort(unique(g)),
        labels = case_when(
          sort(unique(g)) == 2013 ~ "g=2013\nSantiago",
          sort(unique(g)) == 2014 ~ "g=2014\nValdivia",
          sort(unique(g)) == 2017 ~ "g=2017\nAntofagasta",
          TRUE                    ~ paste0("g=", sort(unique(g)))
        ))
    )
}

het_data <- bind_rows(
  tidy_group(cs_high, "High ENCORE"),
  tidy_group(cs_low,  "Low ENCORE")
)

ggplot(het_data, aes(cohort, att, colour = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(
    aes(ymin = ci_lo, ymax = ci_hi, shape = sig),
    position  = position_dodge(0.4),
    linewidth = 0.9,
    size      = 0.8
  ) +
  scale_colour_manual(values = c(col_high, col_low)) +
  scale_shape_manual(values  = c("p<0.05" = 19, "n.s." = 1)) +
  labs(
    title    = "ATT by Cohort and ENCORE Exposure",
    subtitle = "Point estimates with 95% CI | solid = p<0.05",
    x        = "Treatment cohort",
    y        = "ATT (log new credit)",
    colour   = NULL,
    shape    = NULL
  ) +
  theme(legend.position = "bottom")

ggsave(dout("fig6_het_cohort_encore.png"), width = 8, height = 5, dpi = 150)
```



## Summary Comparison Table

```{r comparison}
#| label: comparison

twfe_coef_full <- coef(m4_2)["log1p(lrcases):court"]
twfe_se_full   <- sqrt(diag(vcov(m4_2)))["log1p(lrcases):court"]

cs_h <- aggte(cs_high, type = "simple")
cs_l <- aggte(cs_low,  type = "simple")

comp <- tribble(
  ~Estimator,   ~Sample,         ~ATT,                      ~SE,
  "TWFE",       "Full",          twfe_coef_full,            twfe_se_full,
  "CS (DR)",    "Full",          cs_simple$overall.att,     cs_simple$overall.se,
  "CS (DR)",    "High ENCORE",   cs_h$overall.att,          cs_h$overall.se,
  "CS (DR)",    "Low ENCORE",    cs_l$overall.att,          cs_l$overall.se
) |>
  mutate(
    `ci_lo` = ATT - 1.96 * SE,
    `ci_hi` = ATT + 1.96 * SE,
    across(ATT:`ci_hi`, ~round(., 4))
  )

comp |>
  gt() |>
  tab_header(
    title    = "Estimator Comparison",
    subtitle = "Effect of court introduction × SMA cases on log new credit"
  ) |>
  fmt_number(columns = where(is.numeric), decimals = 4) |>
  cols_label(ci_lo = "CI low", ci_hi = "CI high") |>
  tab_footnote("CS uses doubly-robust method with not-yet-treated controls.")
```

---

# Conclusion {#sec-conclusion}

This document replicates the TWFE analysis of Ossandon Busch et al. (2025) and extends it using two estimators robust to treatment effect heterogeneity across cohorts.

**Main takeaways:**

1. The CS simple ATT is directionally consistent with the TWFE coefficient, suggesting the TWFE results are not severely contaminated by negative weighting.
2. **Cohort heterogeneity** is present: the Antofagasta cohort (g=2017) differs from earlier cohorts, consistent with the very different timing and institutional context of the northern court.
3. The **ENCORE heterogeneity** is preserved under CS — effects are concentrated among high-ENCORE firms — which validates the paper's main identification logic.
4. The **Goodman-Bacon decomposition** reveals the weight structure of TWFE and whether contaminated comparisons drive the result.

---

# Appendix {#sec-appendix}

## Required Data Files

Place the following `.dta` files in the `data/` folder before rendering:

| File | Full path | Used in |
|------|-----------|---------|
| `cmf_project_2011-2024.dta` | `C:/Users/wb631166/OneDrive - WBG/Desktop/CEMLA/CMF/Data/Base Final/` | §1.1 |
| `cmf_project_super_intendencia.dta` | `C:/Users/wb631166/OneDrive - WBG/Desktop/CEMLA/CMF/Data/` | §1.2 |
| `super_IDs.dta` | `C:/Users/wb631166/OneDrive - WBG/Desktop/CEMLA/CMF/Data/` | §1.3 |

All outputs (figures, processed datasets) are saved to `C:/WBG/GitHub/CMF/Data/out/`.

: Required data files {tbl-colwidths="[35, 45, 20]"}

## Required R Packages

```r
install.packages(c(
  "haven", "tidyverse", "fixest", "did",
  "bacondecomp", "modelsummary", "kableExtra",
  "patchwork", "scales", "gt", "DescTools", "sessioninfo"
))
```

## Session Info

```{r session-info}
#| label: session-info
#| code-fold: false

session_info(c("haven", "tidyverse", "fixest", "did",
               "bacondecomp", "modelsummary"))
```
