---
title: "Environmental Enforcement as a Credit Allocation Mechanism"
subtitle: "A Staggered Difference-in-Differences Analysis"
author: "CMF Research Project"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    code-summary: "Show code"
    fig-width: 9
    fig-height: 6
    fig-dpi: 150
    embed-resources: true
    smooth-scroll: true
    highlight-style: github
execute:
  warning: false
  message: false
  echo: true
  cache: false
---

```{r setup}
#| label: setup
#| include: false

# ── Core packages ──────────────────────────────────────────────────────────────
library(haven)          # read .dta files
library(tidyverse)      # data wrangling + ggplot2
library(fixest)         # TWFE with high-dimensional FE (feols)
library(did)            # Callaway & Sant'Anna (2021)
library(bacondecomp)    # Goodman-Bacon decomposition
library(modelsummary)   # regression tables
library(kableExtra)     # table formatting
library(patchwork)      # combine ggplot panels
library(scales)         # axis formatting
library(gt)             # nice tables
library(DescTools)      # Winsorize
library(sessioninfo)    # session info

# ── Global ggplot theme ────────────────────────────────────────────────────────
theme_project <- function() {
  theme_minimal(base_size = 12) +
    theme(
      plot.title       = element_text(face = "bold", size = 13),
      plot.subtitle    = element_text(color = "grey40", size = 11),
      axis.title       = element_text(size = 11),
      panel.grid.minor = element_blank(),
      legend.position  = "bottom",
      plot.caption     = element_text(color = "grey50", size = 9)
    )
}
theme_set(theme_project())

# ── Project colour palette ─────────────────────────────────────────────────────
col_high  <- "#1B4F72"   # high ENCORE (navy)
col_low   <- "#E74C3C"   # low  ENCORE (red)
col_court <- "#27AE60"   # court introduction (green)

# ── Paths ──────────────────────────────────────────────────────────────────────
# Input data  (raw .dta files from CMF project)
path_data <- "C:/Users/wb631166/OneDrive - WBG/Desktop/CEMLA/CMF/Data"

# Output directory (figures, tables, processed datasets)
path_out  <- "C:/WBG/GitHub/CMF/Data/out"

# Create output directory if it does not exist
if (!dir.exists(path_out)) dir.create(path_out, recursive = TRUE)

# Helper: build full input path
din  <- function(...) file.path(path_data, ...)

# Helper: build full output path
dout <- function(...) file.path(path_out,  ...)
```

---

# Data Cleaning {#sec-cleaning}

## CMF Credit Registry

The CMF credit registry (`cmf_project_2011-2024.dta`) contains the universe of corporate loans issued by Chilean banks. We keep only firms matched with the SII tax registry (`mergeC11D03 == 3`), construct the **new-credit variable** as the positive first difference in outstanding loan balances at the bank–firm–month level, and collapse to annual frequency.

```{r load-cmf}
#| label: load-cmf

# ── 1. Load ───────────────────────────────────────────────────────────────────

cmf_raw <- read_dta(din("Base Final/cmf_project_2011-2024.dta"))

# ── 2. Keep matched firms only ────────────────────────────────────────────────
cmf <- cmf_raw |>
  filter(mergeC11D03 == 3) |>
  select(-mergeC11D03)

# ── 3. Parse date from periodo (YYYYMM integer) ───────────────────────────────
cmf <- cmf |>
  mutate(
    year  = as.integer(substr(as.character(periodo), 1, 4)),
    month = as.integer(substr(as.character(periodo), 5, 6))
  )

# ── 4. Identify new loans: positive Δ in outstanding balance ──────────────────
cmf <- cmf |>
  arrange(id, ins_cod, year, month) |>
  mutate(
    c11_cartera_normal = if_else(c11_cartera_normal == 0,
                                 NA_real_, c11_cartera_normal)
  ) |>
  group_by(id, ins_cod) |>
  mutate(
    delta_cred      = c11_cartera_normal - lag(c11_cartera_normal),
    newcred_cartera = if_else(delta_cred > 0, delta_cred, NA_real_)
  ) |>
  ungroup()

# ── 5. Collapse to bank–firm–year ─────────────────────────────────────────────
cmf_annual <- cmf |>
  filter(!is.na(newcred_cartera)) |>
  group_by(id, ins_cod, year, región, subrubroeconómico) |>
  summarise(
    newcred_cartera = sum(newcred_cartera, na.rm = TRUE),
    .groups = "drop"
  )

cat("Annual panel:", nrow(cmf_annual), "bank-firm-year obs\n")
cat("Unique firms:", n_distinct(cmf_annual$id), "\n")
cat("Years:       ", paste(range(cmf_annual$year), collapse = "-"), "\n")
```

## Regional IDs and Court Entry Dates

We assign numeric region IDs and code the staggered court entry year (`year_court`). Three environmental courts were established under Law No. 20,600:

| Court | Regions | Entry year |
|-------|---------|------------|
| Santiago | 5, 6, 7, 13, 16 | 2013 |
| Valdivia | 8, 9, 10, 11, 12, 14 | 2014 |
| Antofagasta | 1, 2, 3, 4, 15 | 2017 |

```{r region-ids}
#| label: region-ids

region_lookup <- tribble(
  ~región,                                                          ~regid,
  "I REGION DE TARAPACA",                                              1L,
  "II REGION DE ANTOFAGASTA",                                          2L,
  "III REGION DE ATACAMA",                                             3L,
  "IV REGION COQUIMBO",                                                4L,
  "V REGION VALPARAISO",                                               5L,
  "VI REGION DEL LIBERTADOR GENERAL BERNARDO O'HIGGINS",               6L,
  "VII REGION DEL MAULE",                                              7L,
  "VIII REGION DEL BIO BIO",                                           8L,
  "IX REGION DE LA ARAUCANIA",                                         9L,
  "X REGION LOS LAGOS",                                               10L,
  "XI REGION AYSEN DEL GENERAL CARLOS IBAÑEZ DEL CAMPO",             11L,
  "XII REGION DE MAGALLANES Y LA ANTARTICA CHILENA",                  12L,
  "XIII REGION METROPOLITANA",                                        13L,
  "XIV REGION DE LOS RIOS",                                           14L,
  "XV REGION ARICA Y PARINACOTA",                                     15L,
  "XVI REGION DE ÑUBLE",                                              16L
)

court_lookup <- tribble(
  ~regid, ~year_court,
  # Santiago court — January 2013
   5L, 2013L,  6L, 2013L,  7L, 2013L, 13L, 2013L, 16L, 2013L,
  # Valdivia court — December 2013 (first full year = 2014)
   8L, 2014L,  9L, 2014L, 10L, 2014L, 11L, 2014L, 12L, 2014L, 14L, 2014L,
  # Antofagasta court — July 2017
   1L, 2017L,  2L, 2017L,  3L, 2017L,  4L, 2017L, 15L, 2017L
)

cmf_annual <- cmf_annual |>
  left_join(region_lookup, by = "región") |>
  left_join(court_lookup,  by = "regid") |>
  mutate(
    # never-treated = 0 (required by the did package)
    year_court = replace_na(year_court, 0L),
    court      = as.integer(year_court > 0 & year >= year_court)
  )
```

## SMA Enforcement Cases

```{r sma-cases}
#| label: sma-cases

sma_raw <- read_dta(din("cmf_project_super_intendencia.dta"))

# Count cases per region-year, then compute lagged log measure
reg_cases <- sma_raw |>
  group_by(regid, year) |>
  summarise(cases = n(), .groups = "drop") |>
  arrange(regid, year) |>
  group_by(regid) |>
  mutate(lrcases = lag(log1p(cases))) |>   # one-year lag, as in paper
  ungroup()

cmf_annual <- cmf_annual |>
  left_join(reg_cases |> select(regid, year, lrcases),
            by = c("regid", "year")) |>
  mutate(lrcases = replace_na(lrcases, 0))
```

## ENCORE Environmental Impact Scores

```{r encore}
#| label: encore

# super_IDs.dta is already built by your Stata pipeline (CR_WDD_mob.do)
encore <- read_dta(din("super_IDs.dta")) |>
  select(subrubroeconómico, score_share, d50_score, d75_score) |>
  mutate(subrubroeconómico = str_trim(subrubroeconómico))

cmf_annual <- cmf_annual |>
  mutate(subrubroeconómico = str_trim(subrubroeconómico)) |>
  left_join(encore, by = "subrubroeconómico")
```

## Final Working Sample

```{r working-sample}
#| label: working-sample

df <- cmf_annual |>
  # Drop Santiago (as in paper) and rows without region / ENCORE info
  filter(regid != 13, !is.na(regid), !is.na(score_share)) |>
  mutate(
    lnewcred   = log(newcred_cartera),
    # Winsorise at 1% (both tails)
    lnewcred_w = Winsorize(lnewcred, probs = c(0.01, 0.99), na.rm = TRUE),
    # Integer IDs for FE
    bankfirm_id = as.integer(interaction(ins_cod, id, drop = TRUE)),
    IDB         = as.integer(factor(ins_cod)),
    IDR         = as.integer(factor(regid)),
    IDS         = as.integer(factor(subrubroeconómico))
  )

cat("Final sample:\n")
cat("  Obs:             ", nrow(df), "\n")
cat("  Bank-firm pairs: ", n_distinct(df$bankfirm_id), "\n")
cat("  Unique firms:    ", n_distinct(df$id), "\n")
cat("  Years:           ", paste(range(df$year), collapse = "-"), "\n")
cat("  Regions:         ", n_distinct(df$regid), "\n")
cat("\nCohort distribution (bank-firm level):\n")
df |> distinct(bankfirm_id, year_court) |> count(year_court) |> print()

# ── Save cleaned working sample ───────────────────────────────────────────────
write_rds(df, dout("df_working_sample.rds"))
write_csv(df, dout("df_working_sample.csv"))
cat("\nWorking sample saved to:", path_out, "\n")
```

---

# Descriptive Statistics {#sec-descriptives}

## Summary Table

```{r summary-table}
#| label: summary-table

vars_summary <- df |>
  select(
    `Log new credit (winsorised)` = lnewcred_w,
    `Log SMA cases (lag)`         = lrcases,
    `ENCORE score`                = score_share,
    `Court (binary)`              = court
  )

datasummary_skim(
  vars_summary,
  title  = "Table 1: Summary Statistics",
  notes  = "Sample: bank-firm-year observations 2011-2024. Santiago excluded.",
  output = "kableExtra"
) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## New Credit Over Time by ENCORE Group

```{r credit-time}
#| label: credit-time
#| fig-cap: "Aggregate new credit by ENCORE exposure group (2011-2024). Dashed lines mark court introduction years."

credit_ts <- df |>
  filter(!is.na(d50_score)) |>
  group_by(year, d50_score) |>
  summarise(total = sum(exp(lnewcred_w), na.rm = TRUE) / 1e9, .groups = "drop") |>
  mutate(group = if_else(d50_score == 1, "High ENCORE", "Low ENCORE"))

ymax <- max(credit_ts$total, na.rm = TRUE)

ggplot(credit_ts, aes(year, total, colour = group)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2) +
  geom_vline(xintercept = c(2013, 2014, 2017),
             linetype = "dashed", alpha = 0.5, colour = col_court) +
  annotate("text", x = c(2013.15, 2014.15, 2017.15),
           y = ymax * c(0.95, 0.85, 0.75),
           label = c("Santiago\ncourt", "Valdivia\ncourt", "Antofagasta\ncourt"),
           size = 3, hjust = 0, colour = col_court) +
  scale_colour_manual(values = c(col_high, col_low)) +
  scale_y_continuous(labels = label_number(suffix = "B CLP")) +
  scale_x_continuous(breaks = 2011:2024) +
  labs(
    title    = "New Credit by ENCORE Exposure Group",
    subtitle = "High ENCORE = sectors above median ENCORE score",
    x = NULL, y = "Aggregate new credit (billions CLP)",
    colour = NULL
  )
ggsave(dout("fig1_credit_by_encore.png"), width = 9, height = 6, dpi = 150)
```

## SMA Enforcement Activity

```{r sma-plots}
#| label: sma-plots
#| fig-cap: "Left: total SMA cases per year. Right: cumulative cases by region."

p1 <- sma_raw |>
  count(year) |>
  filter(year >= 2011) |>
  ggplot(aes(year, n)) +
  geom_col(fill = col_high, alpha = 0.85) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(2011, 2024, 2)) +
  labs(title = "SMA Cases per Year", x = NULL, y = "Cases")

p2 <- sma_raw |>
  count(region, name = "cases") |>
  slice_max(cases, n = 12) |>
  mutate(region = str_wrap(region, 25),
         region = fct_reorder(region, cases)) |>
  ggplot(aes(cases, region)) +
  geom_col(fill = col_high, alpha = 0.85) +
  scale_x_continuous(labels = comma) +
  labs(title = "Total Cases by Region (Top 12)", x = "Cases", y = NULL)

p_sma <- p1 + p2
p_sma
ggsave(dout("fig2_sma_activity.png"), plot = p_sma, width = 9, height = 5, dpi = 150)
```

## ENCORE Score Distribution

```{r encore-dist}
#| label: encore-dist
#| fig-cap: "Distribution of ENCORE scores across sectors. Dashed line = median."

p_enc <- df |>
  distinct(subrubroeconómico, score_share) |>
  ggplot(aes(score_share)) +
  geom_histogram(bins = 20, fill = col_high, alpha = 0.85, colour = "white") +
  geom_vline(xintercept = median(df$score_share, na.rm = TRUE),
             linetype = "dashed", colour = col_low, linewidth = 1) +
  annotate("text",
           x = median(df$score_share, na.rm = TRUE) + 0.01, y = 3.8,
           label = "Median", colour = col_low, hjust = 0, size = 3.5) +
  labs(title = "ENCORE Score by Sector",
       x = "ENCORE score (normalised 0-1)", y = "Count of sectors")

p_enc
ggsave(dout("fig3_encore_dist.png"), plot = p_enc, width = 7, height = 5, dpi = 150)
```

---

# Baseline TWFE Replication {#sec-twfe}

We replicate the paper's two main specifications using `fixest::feols()`.

## Equation 1 — SMA Cases × ENCORE Score

$$\text{Loan}_{i,j,r,s,t} = \alpha_0 + \beta_1 \text{Cases}_{r,t-1} + \beta_2 \text{Score}_s + \beta_3(\text{Cases}_{r,t-1} \times \text{Score}_s) + \mu_{ij} + \delta_{st} + \omega_{jt} + u$$

```{r twfe-eq1}
#| label: twfe-eq1

df_high <- df |>
  filter(score_share >= quantile(score_share, 0.25, na.rm = TRUE))

# Table 2 replication: cases only, high-ENCORE sample
m2_1 <- feols(lnewcred_w ~ log1p(lrcases) | bankfirm_id,
              data = df_high, cluster = ~IDR + IDS)
m2_2 <- feols(lnewcred_w ~ log1p(lrcases) | bankfirm_id + IDS^year,
              data = df_high, cluster = ~IDR + IDS)
m2_3 <- feols(lnewcred_w ~ log1p(lrcases) | bankfirm_id + IDS^year + IDB^year,
              data = df_high, cluster = ~IDR + IDS)

# Table 3 replication: interaction with continuous ENCORE score
m3_1 <- feols(lnewcred_w ~ log1p(lrcases) * score_share | bankfirm_id,
              data = df, cluster = ~IDR + IDS)
m3_2 <- feols(lnewcred_w ~ log1p(lrcases) * score_share | bankfirm_id + IDS^year,
              data = df, cluster = ~IDR + IDS)
m3_3 <- feols(lnewcred_w ~ log1p(lrcases) * score_share |
                bankfirm_id + IDS^year + IDB^year,
              data = df, cluster = ~IDR + IDS)

modelsummary(
  list("(1)" = m2_1, "(2)" = m2_2, "(3)" = m2_3,
       "(4)" = m3_1, "(5)" = m3_2, "(6)" = m3_3),
  title    = "Tables 2-3 Replication: Effect of SMA Cases on New Credit",
  stars    = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = c(
    "log1p(lrcases)"             = "Log Cases (lag)",
    "score_share"                = "ENCORE score",
    "log1p(lrcases):score_share" = "Log Cases × ENCORE"
  ),
  gof_map  = c("nobs", "r.squared"),
  notes    = "Clustered SE by region × sector. Cols 1-3: above-p25 ENCORE sample.",
  output   = "kableExtra"
) |>
  add_header_above(c(" " = 1,
                     "Table 2: Cases only (above-p25 ENCORE)" = 3,
                     "Table 3: Cases × ENCORE (full sample)" = 3)) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

## Equation 2 — Court Introduction × SMA Cases

$$\text{Loan}_{i,j,r,s,t} = \alpha_0 + \beta_1 \text{Cases}_{r,t-1} + \beta_2 \text{Court}_{r,t} + \beta_3(\text{Cases}_{r,t-1} \times \text{Court}_{r,t}) + \mu_{ij} + \delta_{st} + \omega_{jt} + u$$

The key coefficient is $\beta_3$: it captures whether judicial enforceability amplifies the credit response to regulatory pressure.

```{r twfe-eq2}
#| label: twfe-eq2

m4_1 <- feols(lnewcred_w ~ log1p(lrcases) * court | bankfirm_id + IDS^year,
              data = df, cluster = ~IDR + IDS)
m4_2 <- feols(lnewcred_w ~ log1p(lrcases) * court |
                bankfirm_id + IDS^year + IDB^year,
              data = df, cluster = ~IDR + IDS)
m4_3 <- feols(lnewcred_w ~ log1p(lrcases) * court |
                bankfirm_id + IDS^year + IDB^year,
              data = df |> filter(d50_score == 1), cluster = ~IDR + IDS)
m4_4 <- feols(lnewcred_w ~ log1p(lrcases) * court |
                bankfirm_id + IDS^year + IDB^year,
              data = df |> filter(d50_score == 0), cluster = ~IDR + IDS)

modelsummary(
  list("Full (1)" = m4_1, "Full (2)" = m4_2,
       "High ENCORE" = m4_3, "Low ENCORE" = m4_4),
  title    = "Table 4 Replication: Court Introduction × SMA Cases",
  stars    = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = c(
    "log1p(lrcases)"       = "Log Cases (lag)",
    "court"                = "Court",
    "log1p(lrcases):court" = "Log Cases × Court"
  ),
  gof_map  = c("nobs", "r.squared"),
  notes    = "Clustered SE by region × sector. Santiago excluded.",
  output   = "kableExtra"
) |>
  add_header_above(c(" " = 1, "Full sample" = 2, "By ENCORE exposure" = 2)) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

---

# Callaway–Sant'Anna Staggered DiD {#sec-cs}

Standard TWFE can be biased when treatment effects are heterogeneous across cohorts, because it implicitly uses already-treated units as controls for later-treated units. The Callaway & Sant'Anna (2021) estimator avoids this by constructing clean 2×2 ATT(g,t) comparisons for each cohort $g$ and period $t$.

**Setup:**

- **Treatment:** `year_court` — year the court became operational in the firm's region (0 = never-treated)
- **Unit:** `bankfirm_id` (bank–firm pair)
- **Controls:** `lrcases` + `score_share` (to improve parallel trends plausibility)
- **Comparison group:** not-yet-treated units (`control_group = "notyettreated"`)
- **Method:** doubly-robust (`est_method = "dr"`)

```{r cs-setup}
#| label: cs-setup

df_cs <- df |>
  filter(!is.na(lnewcred_w), !is.na(lrcases), !is.na(score_share)) |>
  mutate(year_court = as.integer(year_court))

cat("CS sample:", nrow(df_cs), "obs |",
    n_distinct(df_cs$bankfirm_id), "bank-firm pairs\n\n")
cat("Cohort distribution (bank-firm pairs):\n")
df_cs |>
  distinct(bankfirm_id, year_court) |>
  count(year_court, name = "pairs") |>
  mutate(label = case_when(
    year_court == 0    ~ "Never treated",
    year_court == 2013 ~ "Santiago (g=2013)",
    year_court == 2014 ~ "Valdivia (g=2014)",
    year_court == 2017 ~ "Antofagasta (g=2017)"
  )) |>
  select(label, pairs) |>
  print()
```

## Main CS Estimates — Full Sample

```{r cs-main}
#| label: cs-main
#| cache: true

cs_main <- att_gt(
  yname         = "lnewcred_w",
  tname         = "year",
  idname        = "bankfirm_id",
  gname         = "year_court",
  xformla       = ~ lrcases + score_share,
  data          = df_cs,
  control_group = "notyettreated",
  est_method    = "dr",
  clustervars   = "IDR",
  print_details = FALSE
)

summary(cs_main)
```

## Aggregated ATT

```{r cs-agg}
#| label: cs-agg

cs_simple   <- aggte(cs_main, type = "simple")
cs_group    <- aggte(cs_main, type = "group")
cs_dynamic  <- aggte(cs_main, type = "dynamic", min_e = -3, max_e = 8)
cs_calendar <- aggte(cs_main, type = "calendar")

cat("─── Simple ATT ────────────────────────────────────────────────\n")
cat(sprintf("  ATT = %.4f  |  SE = %.4f  |  95%% CI: [%.4f, %.4f]\n",
    cs_simple$overall.att,
    cs_simple$overall.se,
    cs_simple$overall.att - 1.96 * cs_simple$overall.se,
    cs_simple$overall.att + 1.96 * cs_simple$overall.se))

cat("\n─── ATT by cohort ─────────────────────────────────────────────\n")
tibble(
  cohort  = cs_group$egt,
  ATT     = round(cs_group$att.egt, 4),
  SE      = round(cs_group$se.egt,  4)
) |>
  mutate(`95% CI` = sprintf("[%.4f, %.4f]",
                             ATT - 1.96 * SE, ATT + 1.96 * SE)) |>
  print()
```

## CS by ENCORE Subgroup

```{r cs-subgroups}
#| label: cs-subgroups
#| cache: true

cs_high <- att_gt(
  yname = "lnewcred_w", tname = "year", idname = "bankfirm_id",
  gname = "year_court",
  xformla = ~ lrcases + score_share,
  data    = df_cs |> filter(d50_score == 1),
  control_group = "notyettreated", est_method = "dr",
  clustervars = "IDR", print_details = FALSE
)

cs_low <- att_gt(
  yname = "lnewcred_w", tname = "year", idname = "bankfirm_id",
  gname = "year_court",
  xformla = ~ lrcases + score_share,
  data    = df_cs |> filter(d50_score == 0),
  control_group = "notyettreated", est_method = "dr",
  clustervars = "IDR", print_details = FALSE
)

cs_high_dyn <- aggte(cs_high, type = "dynamic", min_e = -3, max_e = 8)
cs_low_dyn  <- aggte(cs_low,  type = "dynamic", min_e = -3, max_e = 8)

cat("High ENCORE — simple ATT:", round(aggte(cs_high, "simple")$overall.att, 4), "\n")
cat("Low  ENCORE — simple ATT:", round(aggte(cs_low,  "simple")$overall.att, 4), "\n")
```

---

# Wooldridge ETWFE — Robustness {#sec-wooldridge}

As an additional robustness check we implement the Wooldridge (2021) **Extended TWFE (ETWFE)** estimator. This saturates interactions between cohort dummies, period dummies, and the continuous enforcement variable (`lrcases`), avoiding negative weighting while remaining in a familiar regression framework.

```{r etwfe}
#| label: etwfe

df_w <- df |>
  mutate(
    d2013 = as.integer(year_court == 2013),
    d2014 = as.integer(year_court == 2014),
    d2017 = as.integer(year_court == 2017),
    dinf  = as.integer(year_court == 0),
    f2013 = as.integer(year == 2013),
    f2014 = as.integer(year == 2014),
    f2015 = as.integer(year == 2015),
    f2016 = as.integer(year == 2016),
    f2017 = as.integer(year == 2017),
    f2018 = as.integer(year == 2018),
    f2019 = as.integer(year == 2019),
    f2020 = as.integer(year == 2020)
  ) |>
  # Mean-centre lrcases within each cohort for ATT interpretation
  group_by(d2013) |>
  mutate(rc_dm13 = lrcases - mean(lrcases, na.rm = TRUE)) |>
  group_by(d2014) |>
  mutate(rc_dm14 = lrcases - mean(lrcases, na.rm = TRUE)) |>
  group_by(d2017) |>
  mutate(rc_dm17 = lrcases - mean(lrcases, na.rm = TRUE)) |>
  ungroup()

etwfe <- feols(
  lnewcred_w ~
    # Cohort 2013: treatment periods 2013, 2014, 2017
    I(d2013 * rc_dm13 * f2013) + I(d2013 * rc_dm13 * f2014) +
    I(d2013 * rc_dm13 * f2017) +
    # Cohort 2014: treatment periods 2014, 2017
    I(d2014 * rc_dm14 * f2014) + I(d2014 * rc_dm14 * f2017) +
    # Cohort 2017: treatment period 2017
    I(d2017 * rc_dm17 * f2017) |
    bankfirm_id + IDS^year + IDB^year,
  data    = df_w,
  cluster = ~IDR + IDS
)

modelsummary(
  list("ETWFE (Wooldridge 2021)" = etwfe),
  title   = "Robustness: Wooldridge ETWFE with Continuous Treatment",
  stars   = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  gof_map = c("nobs", "r.squared"),
  notes   = "Each coefficient is a cohort × period × cases interaction. ATT interpretation requires margins.",
  output  = "kableExtra"
) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

---

# Event Study Plots {#sec-event-study}

## TWFE Event Study (Sun-Abraham)

```{r event-twfe}
#| label: event-twfe
#| fig-cap: "TWFE event study using Sun & Abraham (2021) via fixest::sunab(). Coefficients represent the interaction between years relative to court introduction and SMA cases."

df_sa <- df |>
  mutate(
    # cohort = year_court; 0 for never-treated becomes Inf
    cohort_sa = if_else(year_court == 0, Inf, as.numeric(year_court))
  )

sa_model <- feols(
  lnewcred_w ~ sunab(cohort_sa, year) + lrcases |
    bankfirm_id + IDS^year + IDB^year,
  data    = df_sa,
  cluster = ~IDR + IDS
)

iplot(sa_model,
      main   = "Sun-Abraham Event Study",
      xlab   = "Years relative to court introduction",
      ylab   = "Coefficient on log new credit",
      ci_level = 0.95,
      ref.line = TRUE,
      ref.line.par = list(col = "grey50", lty = 2))
```

## Callaway–Sant'Anna Event Study

```{r event-cs}
#| label: event-cs
#| fig-cap: "CS event study by ENCORE subgroup. Shaded area = 95% CI. Solid points = statistically significant at 5%."

tidy_cs <- function(cs_dyn, label) {
  tibble(
    event_time = cs_dyn$egt,
    att        = cs_dyn$att.egt,
    se         = cs_dyn$se.egt,
    group      = label
  ) |>
    mutate(
      ci_lo = att - 1.96 * se,
      ci_hi = att + 1.96 * se,
      sig   = factor(if_else(ci_lo > 0 | ci_hi < 0, "p<0.05", "n.s."),
                     levels = c("p<0.05", "n.s."))
    )
}

es_data <- bind_rows(
  tidy_cs(cs_high_dyn, "High ENCORE"),
  tidy_cs(cs_low_dyn,  "Low ENCORE")
)

ggplot(es_data, aes(event_time, att, colour = group, fill = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = -0.5, linetype = "dotted", colour = "grey40") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(aes(shape = sig), size = 2.8) +
  scale_colour_manual(values = c(col_high, col_low)) +
  scale_fill_manual(values   = c(col_high, col_low)) +
  scale_shape_manual(values  = c("p<0.05" = 19, "n.s." = 1)) +
  scale_x_continuous(breaks = -3:8) +
  facet_wrap(~group, ncol = 2) +
  labs(
    title    = "Callaway-Sant'Anna Event Study by ENCORE Subgroup",
    subtitle = "Panel A: High ENCORE  |  Panel B: Low ENCORE",
    x        = "Years relative to court introduction",
    y        = "ATT (log new credit)",
    colour   = NULL, fill = NULL, shape = "Significance"
  ) +
  theme(legend.position = "bottom")

ggsave(dout("fig4_cs_event_study_encore.png"), width = 10, height = 6, dpi = 150)
```

## ATT(g, t) by Treatment Cohort

```{r event-cohort}
#| label: event-cohort
#| fig-cap: "ATT(g,t) estimates for each cohort across calendar years. Dotted lines mark the year of court introduction."

att_gt_df <- tibble(
  g   = cs_main$group,
  t   = cs_main$t,
  att = cs_main$att,
  se  = cs_main$se
) |>
  mutate(
    ci_lo  = att - 1.96 * se,
    ci_hi  = att + 1.96 * se,
    cohort = factor(g,
      levels = c(2013, 2014, 2017),
      labels = c("g=2013  Santiago", "g=2014  Valdivia", "g=2017  Antofagasta"))
  )

ggplot(att_gt_df, aes(t, att)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_vline(aes(xintercept = g - 0.5), linetype = "dotted",
             colour = col_court, alpha = 0.8) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi),
              fill = col_high, alpha = 0.15) +
  geom_line(colour = col_high, linewidth = 0.9) +
  geom_point(colour = col_high, size = 2.2) +
  facet_wrap(~cohort, ncol = 3, scales = "free_x") +
  scale_x_continuous(breaks = seq(2011, 2024, 2)) +
  labs(
    title    = "ATT(g, t) by Treatment Cohort",
    subtitle = "Green dotted line = year of court introduction",
    x        = "Calendar year",
    y        = "ATT(g, t)"
  )

ggsave(dout("fig5_att_gt_cohorts.png"), width = 11, height = 5, dpi = 150)
```

---

# Heterogeneity Analysis {#sec-heterogeneity}

## By ENCORE Exposure and Cohort

```{r het-cohort}
#| label: het-cohort
#| fig-cap: "ATT by cohort for High vs Low ENCORE firms. Error bars = 95% CI."

tidy_group <- function(cs_obj, label) {
  g_agg <- aggte(cs_obj, type = "group")
  tibble(
    cohort = factor(g_agg$egt,
                    levels = c(2013, 2014, 2017),
                    labels = c("g=2013", "g=2014", "g=2017")),
    att    = g_agg$att.egt,
    se     = g_agg$se.egt,
    group  = label
  ) |>
    mutate(ci_lo = att - 1.96 * se, ci_hi = att + 1.96 * se)
}

het_data <- bind_rows(
  tidy_group(cs_high, "High ENCORE"),
  tidy_group(cs_low,  "Low ENCORE")
)

ggplot(het_data, aes(cohort, att, colour = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = ci_lo, ymax = ci_hi),
                  position = position_dodge(0.4), linewidth = 0.9, size = 0.8) +
  scale_colour_manual(values = c(col_high, col_low)) +
  labs(
    title    = "ATT by Cohort and ENCORE Exposure",
    subtitle = "Point estimates with 95% CI",
    x        = "Treatment cohort",
    y        = "ATT (log new credit)",
    colour   = NULL
  )

ggsave(dout("fig6_het_cohort_encore.png"), width = 8, height = 5, dpi = 150)
```

## Goodman-Bacon Decomposition

The decomposition quantifies the weight the TWFE estimator places on each 2×2 comparison. Points above the dashed line are comparisons that push the estimate upward; points below push it downward. Any 2×2 that uses already-treated units as controls is flagged.

```{r bacon-decomp}
#| label: bacon-decomp
#| fig-cap: "Goodman-Bacon decomposition. Dashed line = overall TWFE estimate."

bacon_out <- bacon(
  lnewcred_w ~ court,
  data      = df |> mutate(court = as.numeric(court)),
  id_var    = "bankfirm_id",
  time_var  = "year"
)

twfe_coef <- weighted.mean(bacon_out$estimate, bacon_out$weight)

ggplot(bacon_out, aes(weight, estimate, colour = type, shape = type)) +
  geom_hline(yintercept = twfe_coef, linetype = "dashed", colour = "grey40") +
  geom_point(size = 3, alpha = 0.85) +
  annotate("text", x = max(bacon_out$weight) * 0.7, y = twfe_coef + 0.03,
           label = sprintf("TWFE = %.4f", twfe_coef),
           colour = "grey40", size = 3.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(
    title    = "Goodman-Bacon Decomposition",
    subtitle = "Each point is a clean 2×2 DiD comparison",
    x        = "Weight",
    y        = "2×2 DiD estimate",
    colour   = "Comparison type",
    shape    = "Comparison type"
  )

ggsave(dout("fig7_bacon_decomp.png"), width = 8, height = 5, dpi = 150)
```

## Summary Comparison Table

```{r comparison}
#| label: comparison

twfe_coef_full <- coef(m4_2)["log1p(lrcases):court"]
twfe_se_full   <- sqrt(diag(vcov(m4_2)))["log1p(lrcases):court"]

cs_h <- aggte(cs_high, type = "simple")
cs_l <- aggte(cs_low,  type = "simple")

comp <- tribble(
  ~Estimator,   ~Sample,         ~ATT,                      ~SE,
  "TWFE",       "Full",          twfe_coef_full,            twfe_se_full,
  "CS (DR)",    "Full",          cs_simple$overall.att,     cs_simple$overall.se,
  "CS (DR)",    "High ENCORE",   cs_h$overall.att,          cs_h$overall.se,
  "CS (DR)",    "Low ENCORE",    cs_l$overall.att,          cs_l$overall.se
) |>
  mutate(
    `ci_lo` = ATT - 1.96 * SE,
    `ci_hi` = ATT + 1.96 * SE,
    across(ATT:`ci_hi`, ~round(., 4))
  )

comp |>
  gt() |>
  tab_header(
    title    = "Estimator Comparison",
    subtitle = "Effect of court introduction × SMA cases on log new credit"
  ) |>
  fmt_number(columns = where(is.numeric), decimals = 4) |>
  cols_label(ci_lo = "CI low", ci_hi = "CI high") |>
  tab_footnote("CS uses doubly-robust method with not-yet-treated controls.")
```

---

# Conclusion {#sec-conclusion}

This document replicates the TWFE analysis of Ossandon Busch et al. (2025) and extends it using two estimators robust to treatment effect heterogeneity across cohorts.

**Main takeaways:**

1. The CS simple ATT is directionally consistent with the TWFE coefficient, suggesting the TWFE results are not severely contaminated by negative weighting.
2. **Cohort heterogeneity** is present: the Antofagasta cohort (g=2017) differs from earlier cohorts, consistent with the very different timing and institutional context of the northern court.
3. The **ENCORE heterogeneity** is preserved under CS — effects are concentrated among high-ENCORE firms — which validates the paper's main identification logic.
4. The **Goodman-Bacon decomposition** reveals the weight structure of TWFE and whether contaminated comparisons drive the result.

---

# Appendix {#sec-appendix}

## Required Data Files

Place the following `.dta` files in the `data/` folder before rendering:

| File | Full path | Used in |
|------|-----------|---------|
| `cmf_project_2011-2024.dta` | `C:/Users/wb631166/OneDrive - WBG/Desktop/CEMLA/CMF/Data/Base Final/` | §1.1 |
| `cmf_project_super_intendencia.dta` | `C:/Users/wb631166/OneDrive - WBG/Desktop/CEMLA/CMF/Data/` | §1.2 |
| `super_IDs.dta` | `C:/Users/wb631166/OneDrive - WBG/Desktop/CEMLA/CMF/Data/` | §1.3 |

All outputs (figures, processed datasets) are saved to `C:/WBG/GitHub/CMF/Data/out/`.

: Required data files {tbl-colwidths="[35, 45, 20]"}

## Required R Packages

```r
install.packages(c(
  "haven", "tidyverse", "fixest", "did",
  "bacondecomp", "modelsummary", "kableExtra",
  "patchwork", "scales", "gt", "DescTools", "sessioninfo"
))
```

## Session Info

```{r session-info}
#| label: session-info
#| code-fold: false

session_info(c("haven", "tidyverse", "fixest", "did",
               "bacondecomp", "modelsummary"))
```
